---
title: "lab10"
author: "Yuwei Wu"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
dbListTables(con)
```


TIP: Use can use the following QUERY to see the structure of a table
```{sql, connection=con}
PRAGMA table_info(actor)
```


## Exercise 1
Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.
```{r}
dbGetQuery(con, "
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name 
LIMIT 15")
```
Try in SQL directly:

```{sql, connection=con}
SELECT   actor_id, first_name, last_name
FROM    actor
ORDER by last_name, first_name
LIMIT 5
```

## Exercise 2
Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.
```{r}
dbGetQuery(con, "
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
ORDER BY last_name
")
```

## Exercise 3
Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.
```{r}
dbGetQuery(con, "
SELECT DISTINCT customer_id
FROM rental
WHERE date(rental_date) = '2005-07-05'
GROUP BY customer_id
")
```

## Exercise 4
### Exercise 4.1
Construct a query that retrives all rows from the payment table where the amount is either 1.99, 7.99, 9.99.
```{r}
dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)
LIMIT 10
")
```

Exercise 4.2
Construct a query that retrives all rows from the payment table where the amount is greater then 5
```{r}
dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount > 5
LIMIT 10
")
```

### Exercise 4.3

Construct a query that retrieves all rows from the payment table where the amount is greater then 5 and less then 8
```{r}
dbGetQuery(con, "
SELECT *
FROM payment
WHERE amount > 5 AND amount > 8
GROUP BY payment_id
LIMIT 10
")
```

## Exercise 5
Retrieve all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.
```{r}
dbGetQuery(con, "
SELECT payment_id, amount
FROM payment AS p INNER JOIN customer AS c   ON p.customer_id = c.customer_id
WHERE c.last_name = 'DAVIS'
")
```

## Exercise 6
### Exercise 6.1
Use COUNT(*) to count the number of rows in rental
```{r}
dbGetQuery(con, "
SELECT COUNT(*)
FROM rental
")
```

### Exercise 6.2
Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id
```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS 'N Rentals'
FROM rental
GROUP BY customer_id
LIMIT 8
")
```

## Exercise 6.3
Repeat the previous query and sort by the count in descending order
```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS 'N Rentals'
FROM rental
GROUP BY customer_id
ORDER BY `N Rentals` DESC
/*ORDER BY - `N Rentals`*/
LIMIT 8
")
```

## Exercise 6.4
Repeat the previous query but use HAVING to only keep the groups with 40 or more.
```{r}
dbGetQuery(con, "
SELECT customer_id, COUNT(*) AS N
FROM rental
GROUP BY customer_id
HAVING N >= 40 
ORDER BY N DESC
")
```

# Exercise 7
The following query calculates a number of summary statistics for the payment table using MAX, MIN, AVG and SUM
```{r}
dbGetQuery(con, "
SELECT
  MAX(amount) AS max,
  MIN(amount) AS min,
  AVG(amount) AS average,
  SUM(amount) as sum
FROM payment
")
```

## Exercise 7.1
Modify the above query to do those calculations for each customer_id
```{r}
dbGetQuery(con, "
SELECT
  customer_id,
  COUNT(*) AS N,
  MAX(amount) AS max,
  MIN(amount) AS min,
  AVG(amount) AS average,
  SUM(amount) as sum
FROM payment
GROUP BY customer_id
LIMIT 5
")
```

## Exercise 7.2
Modify the above query to only keep the customer_ids that have more then 5 payments
```{r}
dbGetQuery(con, "
SELECT
  customer_id,
  COUNT(*) AS N,
  MAX(amount) AS max,
  MIN(amount) AS min,
  AVG(amount) AS average,
  SUM(amount) as sum
FROM payment
GROUP BY customer_id
HAVING N > 5
")
```

## Cleanup
Run the following chunk to disconnect from the connection.
```{r}
# clean up
dbDisconnect(con)
```