---
title: "Lab4"
author: "Yuwei Wu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(lubridate)
```

## 1. Read in the data
```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("met_all.gz")

library(tidyverse)
library(data.table)
library(lubridate)
```
## 2. Prepare the data
Remove temperatures less than -17C
```{r}
met <- met[temp>-17][elev==9999.0,elve := NA]
```
Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).

```{r creat-ymd, cache = TRUE}
met<- met[ , ymd:= as.Date(paste(year, month, day, sep = "-"))]
```

Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[ , table(week(ymd))]
met<- met[week(ymd)==31]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
```{r}
met[, .(
    temp = max(temp,na.rm = T),
    rh = max(rh,na.rm = T),
    wind.sp = max(wind.sp,na.rm = T),
    vis.dist = max(vis.dist,na.rm = T),
    dew.point = max(dew.point,na.rm = T),
    lat = max(lat,na.rm = T),
    lon = max(lon,na.rm = T),
    elev = max(elev,na.rm = T)
), by = "USAFID"]

```
```{r check-avg}
met_avg <-met[, .(
    temp = mean(temp,na.rm = T),
    rh = mean(rh,na.rm = T),
    wind.sp = mean(wind.sp,na.rm = T),
    vis.dist = mean(vis.dist,na.rm = T),
    dew.point = mean(dew.point,na.rm = T),
    lat = mean(lat,na.rm = T),
    lon = mean(lon,na.rm = T),
    elev = mean(elev,na.rm = T)
), by = "USAFID"]
```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
```{r}
met_avg<-met_avg[, region := fifelse(lon > - 98 & lat > 39.71, "NE",
                fifelse(lon < - 98 & lat > 39.71, "NW",
                fifelse(lon < - 98 & lat <= 39.71, "SW","SE")))
     ]
table(met_avg$region)
```


Create a categorical variable for elevation as in the lecture slides
```{r}
met_avg[, elev.cat := fifelse(elev>252, "high", "low")]
```

## 3. Use geom_violin to examine the wind speed and dew point temperature by region
Use facets
Make sure to deal with NA category
Describe what you observe in the graph

```{r}
met_avg[!is.na(region)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = dew.point, color=region, fill=region)) + 
  facet_wrap(~ region, nrow = 1)
```
```{r}
met_avg[!is.na(region)& !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = wind.sp, color=region, fill=region)) + 
  facet_wrap(~ region, nrow = 1)
```
```{r}
met_avg[!is.na(region)& !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = wind.sp, color=region, fill=region)) + 
  facet_wrap(~ region, nrow = 2)
```

The highest dew point temperatures are reported in the southeast.



## 4. Use geom_jitter with stat_smooth to examine the association between dew point temperature and wind speed by region
Colour points by region
Make sure to deal with NA category
Fit a linear regression line by region
Describe what you observe in the graph

```{r}
met_avg[!is.na(region)& !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp,y = dew.point)) + 
  geom_point(mapping = aes(color = region)) + 
  geom_smooth(mapping = aes(linetype = region))
```


```{r}
met_avg[!is.na(region)& !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp,y = dew.point)) + 
  geom_point(mapping = aes(color = region)) + 
  geom_smooth(method = lm, mapping = aes(linetype = region))+
  facet_wrap(~ region, nrow = 2)
```
## 5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region
```{r}

```
## 6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars
```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x=region, y=dew.point)) + 
    stat_summary(fun.data = mean_sdl, geom = "errorbar")
```
I can show with means or with error bars, but I'd like to show both.

## 7. Make a map showing the spatial trend in relative h in the US
```{r}
library(leaflet)
# Generating a color palette
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal
```

Use addMarkers to include the top 10 places in relative h (hint: this will be useful rank(-rh) <= 10)

```{r}
top10rh <- met_avg[ rank(-rh) <= 10]
```


```{r}
rhmap <- leaflet(met_avg) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humid.', opacity=1)
rhmap
```

