---
title: "lab6"
author: "Yuwei Wu"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidytext)
library(dtplyr)
library(ggplot2)
library(forcats)
```

## 1. Read in the data
```{r read-data, cache= TRUE}
if (!file.exists("mtsamples.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    destfile = "mtsamples.csv",
    method   = "libcurl",
    timeout  = 60
    )
mts <- read.csv("mtsamples.csv")
str(mts)
mts<- as_tibble(mts)
```


## What specialties do we have? What are the frequencies of each?
```{r specialties, cache=TRUE}
library(dplyr)
specialties <- count(mts, medical_specialty)
specialties %>%
    arrange(desc(n)) %>% 
      knitr::kable()
```
There are `r nrow(specialties)` medical specialties. The frequency is not evenly distributed. Surgery is most common with 1103 entries, Hospice - Palliative Care is the least frequent with 6 entries. Some are related - ie Consult - History and Physical and Discharge Summary are not specialty-specific. 
```{r graph specialties}
library(dplyr)
library(ggplot2)
library(forcats)
specialties %>% 
  top_n(10) %>%
  ggplot(aes(x=n, y= fct_reorder(medical_specialty, n))) +
  geom_col()
```
## Tokenize words in transcription and look at top 20
```{r tokenize, cache = TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort=TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(x=n, y=fct_reorder(word, n))) +
  geom_col()
```
There are a lot of stop words here. The top word is "the" while we do see "patient". We should get rid of the stop words.

## Remove stop words
```{r remove stopwords, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort=TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  filter(!grepl(pattern = "^[0-9]+$", x=word)) %>%
  top_n(20, n) %>%
  ggplot(aes(x=n, y=fct_reorder(word, n))) +
  geom_col()
  
```
These look more likel medical terms. Numbers were removed from the list of words.

## Repeat question 2, but this time tokenize into bi-grams. how does the result change if you look at tri-grams?
```{r bigrams, cache=TRUE}
mts %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  count(bigram, sort=TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(x=n, y=fct_reorder(bigram, n))) +
  geom_col()
```
"The patient" was the top bigram. There are a lot of non-interesting bigrams (of the, in the).

```{r trigrams, cache=TRUE}
mts %>%
  unnest_ngrams(trigram, transcription, n=3) %>%
  count(trigram, sort=TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(x=n, y=fct_reorder(trigram, n))) +
  geom_col()
```
Top 20 trigrams seem more medical.

## Using the results you got from questions 4. Pick a word and count the words that appears after and before it.
```{r bigrams next word, cache=TRUE}
ptbigram <- 
  mts %>%
  unnest_ngrams(bigram, transcription, n=2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  select(word1, word2) %>%
  filter(word1 == "patient" | word2 == "patient")
```

Words appearing before patient
```{r bigrams before patient, cache=TRUE}
ptbigram %>%
  filter(word2 == "patient") %>%
  count(word1, sort=TRUE) %>%
  anti_join(stop_words, by = c("word1" = "word")) %>%
  top_n(10, n) %>%
  ggplot(aes(x=n, y=fct_reorder(word1, n))) +
  geom_col()
```
Words appearing after patient
```{r bigrams after patient, cache=TRUE}
ptbigram %>%
  filter(word1 == "patient") %>%
  count(word2, sort=TRUE) %>%
  anti_join(stop_words, by = c("word2" = "word")) %>%
  top_n(10, n) %>%
  ggplot(aes(x=n, y=fct_reorder(word2, n))) +
  geom_col()
```
Looks reasonable!

## Which words are most used in each of the specialties. you can use group_by() and top_n() from dplyr to have the calculations be done within each specialty. Remember to remove stopwords. How about the most 5 used words?
```{r most common by specialty, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  group_by(medical_specialty) %>%
  count(word, sort=TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  filter(!grepl(pattern = "^[0-9]+$", x=word)) %>%
  top_n(5, n) %>%
  arrange(medical_specialty, desc(n)) %>%
knitr::kable()
```


